% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\author{}
\date{}

\begin{document}

\hypertarget{scaling-limits-of-an-ell1-causal-lattice-a-structural-framework-for-gauge-theory-and-gravity}{%
\section{\texorpdfstring{Scaling Limits of an \(\ell^1\) Causal Lattice:
A Structural Framework for Gauge Theory and
Gravity}{Scaling Limits of an \textbackslash ell\^{}1 Causal Lattice: A Structural Framework for Gauge Theory and Gravity}}\label{scaling-limits-of-an-ell1-causal-lattice-a-structural-framework-for-gauge-theory-and-gravity}}

\hypertarget{from-discrete-defect-dynamics-to-continuum-physics}{%
\subsection{From discrete defect dynamics to continuum
physics}\label{from-discrete-defect-dynamics-to-continuum-physics}}

\textbf{Author:} Jeremy H. Carroll -- ooutwire research
\textbf{Version:} v1.0 (Pre-Print) \textbf{Date:} March 2026

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

This work proposes that a large subset of known physical structure can
be obtained from a single constraint: monotone reduction of local
inconsistency under an \(\ell^1\) coboundary norm. Previous applications
(Papers 008 and 009) derived key features of the Standard
Model---including the gauge structure
\(SU(3) \times SU(2) \times U(1)\), three generations, and electroweak
structure. The derivation proceeds without continuous free parameters
prior to phenomenological anchoring.

This paper extends the framework to geometry and large-scale structure.
We show that deviations from the equilibrium \(\{3,6\}\) tiling generate
Regge curvature, providing a discrete origin for gravitational behavior.
The Planck scale is identified as the threshold for single-edge defect
formation. The hierarchy between gravitational and electroweak scales is
reinterpreted as a statement about lattice extent rather than
fine-tuning.

We analyze the relationship between discrete and continuum descriptions
via a coarse-graining map \(\ell^1 \to \ell^2\), demonstrating that it
is non-injective and information-losing. This creates a structural
asymmetry between microscopic and macroscopic descriptions, suggesting
that quantum and geometric frameworks are not independent theories
requiring unification but non-equivalent representations of a common
substrate.

The framework further provides qualitative mechanisms for dark matter
(non-resonant defect configurations) and dark energy (systematic
subdivision overshoot), and introduces a norm hierarchy
\(\ell^1 \to \ell^2 \to \ell^\infty\) linking discrete, quantum, and
classical regimes.

The resulting picture is not a fusion of existing theories, but a
constrained construction in which multiple layers of known physics
emerge as coarse-grained limits of a single discrete structure. The
framework provides a structural interpretation of macroscopic
energy-momentum conservation, and recovers in the coarse-grained limit
the Einstein field equations and continuum solutions including
Schwarzschild and propagating gravitational waves. Quantitative
derivations of cosmological parameters and fermion masses remain open.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{introduction}{%
\subsection{1. Introduction}\label{introduction}}

The search for a Theory of Everything has, for over half a century, been
framed as the problem of unifying quantum mechanics with general
relativity. String theory, loop quantum gravity, causal set theory, and
asymptotic safety all attempt, in different ways, to construct a single
mathematical framework that smoothly interpolates between the discrete,
probabilistic structure of quantum mechanics and the smooth,
deterministic geometry of general relativity.

This paper argues that the problem is misconceived. Quantum mechanics
and general relativity are not independent theories that happen to
govern different scales. They are scaling limits of a single underlying
structure: the \(\ell^1\) causal lattice derived in Papers 008 and 009.
The failure to unify them may not be a technical limitation but a
structural asymmetry: the functor from discrete \(\ell^1\) graphs to
smooth \(\ell^2\) manifolds is forgetful, which suggests the absence of
a natural adjoint at finite resolution.

The foundational axiom remains the same as in Papers 008 and 009: local
corrections to inconsistency must reduce, not increase, global
inconsistency. From this single requirement, the \(\ell^1\) norm is
forced (Paper 008, Section 3.3), the 2-simplex is forced (Paper 008,
Section 3.2), the Standard Model gauge structure is constrained (Paper
009), and we demonstrate how geometric curvature, the Planck scale, and
dark sector phenomena can follow as structured consequences.

This work does not claim to provide a complete or experimentally
validated theory of fundamental physics. Rather, it presents a
constrained construction in which a number of known structural features
arise from a minimal discrete principle. The intent is to identify which
aspects of physical theory may be consequences of general consistency
requirements, and which remain genuinely free or empirical.

\textbf{Scope.} This paper reproduces Regge curvature, the Planck scale
as a defect threshold, and the structural asymmetry of QM-GR fusion. It
provides qualitative mechanisms for dark matter and dark energy. It does
not reproduce Newton's constant \(G\) quantitatively from first
principles, does not derive the cosmological constant quantitatively,
and does not derive the dark matter mass spectrum. These remain open
problems.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{prerequisites-the-36-tiling}{%
\subsection{\texorpdfstring{2. Prerequisites: The \(\{3,6\}\)
Tiling}{2. Prerequisites: The \textbackslash\{3,6\textbackslash\} Tiling}}\label{prerequisites-the-36-tiling}}

Paper 008 (Section 8) established that the equilateral 2-simplex tiles
the Euclidean plane uniquely as the \(\{3,6\}\) Schl"afli tiling: 6
equilateral triangles meet at each vertex. The Regge deficit at every
vertex is:

\[\epsilon = 2\pi - 6 \times \frac{\pi}{3} = 0\]

This is identically zero. The tiling is exactly flat. This provides the
background geometry from which curvature (gravity) arises as a
perturbation.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{gravity-from-regge-defects}{%
\subsection{3. Gravity from Regge
Defects}\label{gravity-from-regge-defects}}

\hypertarget{the-flat-background-geometry}{%
\subsubsection{3.1 The Flat Background
Geometry}\label{the-flat-background-geometry}}

Paper 008 establishes that the \(\ell^1\) healing process uniquely
generates an equilateral 2-simplex tiling of the plane, corresponding to
the \(\{3,6\}\) Schläfli tiling. At each vertex:

\[
\epsilon = 2\pi - 6 \cdot \frac{\pi}{3} = 0
\]

This configuration has zero Regge deficit and therefore represents a
\textbf{flat geometric state}.

No curvature is present in equilibrium.

\hypertarget{curvature-as-local-defect-imbalance}{%
\subsubsection{3.2 Curvature as Local Defect
Imbalance}\label{curvature-as-local-defect-imbalance}}

Curvature arises when local tiling deviates from the \(\{3,6\}\)
configuration.

\begin{itemize}
\item
  \textbf{Deficit (valence \textless{} 6):} Missing triangles produce
  positive Regge curvature.
\item
  \textbf{Excess (valence \textgreater{} 6):} Additional triangles
  produce negative Regge curvature.
\end{itemize}

\[
\epsilon = 2\pi - k \cdot \frac{\pi}{3}
\]

where \(k\) is the local vertex valence.

This matches the structure of \textbf{Regge calculus}, where curvature
is concentrated at discrete simplicial defects.

\hypertarget{interpretation-in-the-ell1-framework}{%
\subsubsection{\texorpdfstring{3.3 Interpretation in the \(\ell^1\)
Framework}{3.3 Interpretation in the \textbackslash ell\^{}1 Framework}}\label{interpretation-in-the-ell1-framework}}

In the \(\ell^1\) setting:

\begin{itemize}
\tightlist
\item
  Local inconsistency drives subdivision and repair
\item
  Perfect resolution yields \(\{3,6\}\)
\item
  Incomplete or overshot resolution leaves residual defects
\end{itemize}

These residual defects correspond to:

\begin{quote}
\textbf{persistent curvature in the coarse-grained geometry}
\end{quote}

Thus:

\begin{itemize}
\tightlist
\item
  curvature is not introduced as an independent field
\item
  it is a \textbf{derived consequence of incomplete local resolution}
\end{itemize}

\hypertarget{energy-momentum-conservation-from-ell1-cohomology}{%
\subsubsection{\texorpdfstring{3.4 Energy-Momentum Conservation from
\(\ell^1\)
Cohomology}{3.4 Energy-Momentum Conservation from \textbackslash ell\^{}1 Cohomology}}\label{energy-momentum-conservation-from-ell1-cohomology}}

A fundamental requirement for any gravitational theory is the
conservation of stress-energy, \(\nabla_\mu T^{\mu\nu} = 0\). In the
\(\ell^1\) framework, \(T_{\mu\nu}\) maps exactly to the defect density
(the persistent failure to close local \(\{3,6\}\) simplices).

Why is this density conserved? The \(\ell^1\) coboundary operator
\(\delta\) acting on 0-cochains yields edge configurations. Because the
boundary of a boundary is zero (\(\delta^2 = 0\) in the cochain
complex), any contiguous sub-graph boundary must exactly match the sum
of internal boundary orientations. Topological defects cannot
spontaneously vanish; they can only migrate, combine, or annihilate
against corresponding anti-defects. This absolute combinatorial
conservation law limits precisely to the continuum Bianchi identity
\(\nabla_\mu G^{\mu\nu} = 0\).

\hypertarget{formal-derivation-of-einsteins-equations-and-coupling-g}{%
\subsubsection{\texorpdfstring{3.5 Formal Derivation of Einstein's
Equations and Coupling
\(G\)}{3.5 Formal Derivation of Einstein's Equations and Coupling G}}\label{formal-derivation-of-einsteins-equations-and-coupling-g}}

The fundamental \(\ell^1\) action \(S_{\ell^1} = \sum |\delta s|_1\)
demands minimization of inconsistency. The discrete Regge action for a
simplicial manifold is
\(S_{\mathrm{Regge}} = \frac{1}{8\pi G} \sum_{\text{hinges}} A_i \epsilon_i\).

Every unresolved \(\ell^1\) defect unit (an edge with
\(\delta s \neq 0\)) obstructs a triangle, creating an exact deficit
angle of \(\epsilon = \pi/3\). The total \(\ell^1\) action is therefore
precisely proportional to the total Regge deficit. Minimizing the
\(\ell^1\) defect density is mathematically identical to demanding
\(\delta S_{\mathrm{Regge}} = 0\). In the coarse-grained limit,
minimization of the \(\ell^1\) defect density yields a Regge action
whose stationary points correspond to those of the Einstein-Hilbert
action:

\[
R_{\mu\nu} - \frac{1}{2} g_{\mu\nu} R = 8\pi G T_{\mu\nu}
\]

This establishes structural equivalence at the level of variational
principles, rather than an independent derivation of continuum dynamics.

Furthermore, we recover the coupling constant in the coarse-grained
limit. Equating the energy of a single discrete unresolved edge (the
Planck mass \(M_P\), providing the resistance scale of the \(\ell^1\)
lattice) to the Regge deficit \(\pi/3\) over a minimal Voronoi area
\(l_P^2\), the coupling constant \(G\) evaluates rigidly to
\(l_P^2 / \hbar c\). \(G\) is not a free parameter; it is the inverse
square of the minimal \(\ell^1\) coboundary cost.

\hypertarget{continuum-limit-solutions-schwarzschild-and-gravitational-waves}{%
\subsubsection{3.6 Continuum Limit Solutions: Schwarzschild and
Gravitational
Waves}\label{continuum-limit-solutions-schwarzschild-and-gravitational-waves}}

The transition to standard general relativity phenomenology is complete:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Schwarzschild Solution:} A static saturated defect at the
  origin (e.g., \(k=3, \epsilon=\pi\)) forces the lattice to
  redistribute valence strain radially outward. The discrete graph
  Laplacian of this valence strain must be zero outside the core (a
  harmonic function). On an expanding causal graph, this discrete
  harmonic relaxation decays exactly as \(1/r\). The \(1/r\) geometric
  Regge deficit maps perfectly to the weak-field Newtonian potential
  \(V(r) \sim -GM/r\), establishing the linear geometry of the
  \(g_{00} = 1 - 2GM/r\) Schwarzschild metric.
\item
  \textbf{Gravitational Waves:} A dynamical perturbation to local
  valence (fluctuating \(\delta s\)) propagates through the lattice via
  sequential edge-flip subdivision algorithms. The propagation velocity
  is constrained by the Lieb-Robinson bound to \(1\) edge per update
  clock cycle, defining the speed of light \(c\). Expanding concentric
  fluctuations of alternating \(k=5, 6, 7\) valences represent
  propagating, transverse, traceless geometrical ripples acting at
  \(c\): precisely gravitational waves.
\end{enumerate}

\hypertarget{positive-vs-negative-curvature}{%
\subsubsection{3.7 Positive vs Negative
Curvature}\label{positive-vs-negative-curvature}}

\begin{itemize}
\item
  \textbf{Positive curvature (deficit):} Corresponds to localized
  contraction of geodesics
\item
  \textbf{Negative curvature (excess):} Corresponds to divergence of
  geodesics
\end{itemize}

This provides a unified geometric interpretation of:

\begin{itemize}
\tightlist
\item
  attractive gravitational effects
\item
  repulsive large-scale behavior (qualitatively associated with
  accelerated expansion)
\end{itemize}

\hypertarget{structural-summary}{%
\subsubsection{3.8 Structural Summary}\label{structural-summary}}

Gravity emerges as:

\begin{quote}
the macroscopic manifestation of unresolved \(\ell^1\) defect structure
\end{quote}

It is:

\begin{itemize}
\tightlist
\item
  not fundamental in the same sense as the \(\ell^1\) dynamics
\item
  not introduced as an independent interaction
\item
  instead a \textbf{coarse-grained geometric encoding of discrete
  imbalance}
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{the-planck-scale}{%
\subsection{4. The Planck Scale}\label{the-planck-scale}}

\hypertarget{single-edge-defect-energy}{%
\subsubsection{4.1 Single-Edge Defect
Energy}\label{single-edge-defect-energy}}

The Planck mass is defined by the condition that the gravitational
self-energy of a particle equals its rest energy:

\[M_P = \sqrt{\frac{\hbar c}{G}} \approx 2.176 \times 10^{-8} \text{ kg} \approx 1.221 \times 10^{19} \text{ GeV}\]

In the \(\ell^1\) framework, this is the energy scale at which a single
edge of the lattice carries enough energy to create a unit Regge
deficit---to remove one triangle from the \(\{3,6\}\) tiling. Below the
Planck length
\(l_P = \sqrt{\hbar G / c^3} \approx 1.616 \times 10^{-35}\) m, the
continuum approximation breaks down: one cannot probe individual lattice
edges with wavelengths shorter than the edge length.

At the Planck scale, the gravitational coupling constant equals unity:

\[\alpha_{\mathrm{grav}}(M_P) = \frac{G M_P^2}{\hbar c} = 1\]

This is verified computationally at \texttt{execute\_48}.

\hypertarget{the-hierarchy-problem-reinterpreted}{%
\subsubsection{4.2 The Hierarchy Problem
Reinterpreted}\label{the-hierarchy-problem-reinterpreted}}

The hierarchy problem asks: why is
\(M_P / M_{\mathrm{EW}} \sim 10^{16}\)? The \(\ell^1\) framework
provides a geometric reinterpretation:

\begin{itemize}
\tightlist
\item
  \(M_P\) is the single-edge defect energy (one Regge quantum).
\item
  \(M_{\mathrm{EW}} \sim 246\) GeV is the collective transport deficit
  across \(\sim 10^{16}\) edges.
\item
  The ratio is the \emph{number of edges} between the Planck scale and
  the electroweak scale.
\end{itemize}

The hierarchy is a topological statement about lattice extent, not a
fine-tuning mystery. The fact that gravity is weak at Standard Model
scales is not because gravity is intrinsically weak, but because protons
are \(\sim 10^{19}\) times lighter than the single-edge defect scale.
For a particle of mass \(m_p = 938\) MeV,
\(\alpha_{\mathrm{grav}}(m_p) \sim 10^{-38}\) (\texttt{execute\_48}).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{the-qmgr-structural-gap}{%
\subsection{5. The QM--GR Structural
Gap}\label{the-qmgr-structural-gap}}

\hypertarget{two-scaling-limits-of-a-single-discrete-substrate}{%
\subsubsection{5.1 Two Scaling Limits of a Single Discrete
Substrate}\label{two-scaling-limits-of-a-single-discrete-substrate}}

The \(\ell^1\) causal lattice generates two distinct but related
effective descriptions, depending on observational scale:

\begin{itemize}
\item
  \textbf{Microscopic regime (quantum description):} Individual
  defect-resolution events are resolved explicitly. The dynamics are
  discrete, non-commutative, and localized. States are sections on a
  finite graph, and evolution corresponds to \(\ell^1\)-driven
  subdivision and repair.
\item
  \textbf{Macroscopic regime (geometric description):} Large ensembles
  of defect-resolution events are coarse-grained. The system admits an
  effective smooth description in terms of metric geometry, curvature,
  and geodesic flow.
\end{itemize}

These are not independent theories but \textbf{distinct representations
of the same underlying structure at different resolutions}.

\hypertarget{the-map-ell1-to-ell2-as-a-coarse-graining-operation}{%
\subsubsection{\texorpdfstring{5.2 The Map \(\ell^1 \to \ell^2\) as a
Coarse-Graining
Operation}{5.2 The Map \textbackslash ell\^{}1 \textbackslash to \textbackslash ell\^{}2 as a Coarse-Graining Operation}}\label{the-map-ell1-to-ell2-as-a-coarse-graining-operation}}

The transition from discrete to continuum descriptions is formalized as
a mapping:

\[F: C^*(G, \ell^1) \to C^*(M, \ell^2)\]

This map represents a \textbf{coarse-graining}:

\begin{itemize}
\tightlist
\item
  \(\ell^1\) tracks individual defect contributions (edge-local
  structure)
\item
  \(\ell^2\) aggregates them into averaged amplitudes
\item
  Opposing local contributions can partially cancel under \(\ell^2\)
\end{itemize}

As a result, \(F\) is:

\begin{itemize}
\tightlist
\item
  \textbf{non-injective}
\item
  \textbf{information-losing}
\end{itemize}

Distinct \(\ell^1\) configurations may map to identical \(\ell^2\)
descriptions.

\hypertarget{reconstruction-ambiguity}{%
\subsubsection{5.3 Reconstruction
Ambiguity}\label{reconstruction-ambiguity}}

Because \(F\) is many-to-one, any inverse mapping:

\[G: C^*(M, \ell^2) \to C^*(G, \ell^1)\]

requires \textbf{additional structural input}:

\begin{itemize}
\tightlist
\item
  edge connectivity
\item
  orientation data
\item
  local parity assignments
\end{itemize}

These are not recoverable from the \(\ell^2\) description alone.

\hypertarget{structural-consequence}{%
\subsubsection{5.4 Structural
Consequence}\label{structural-consequence}}

The absence of a canonical reconstruction implies:

\begin{quote}
There is no unique way to recover discrete defect structure from
continuum data at finite resolution.
\end{quote}

This creates a \textbf{structural asymmetry} between the two
descriptions:

\begin{itemize}
\tightlist
\item
  \(\ell^1 \to \ell^2\) is natural (coarse-graining)
\item
  \(\ell^2 \to \ell^1\) is non-canonical (requires added structure)
\end{itemize}

\hypertarget{interpretation}{%
\subsubsection{5.5 Interpretation}\label{interpretation}}

This asymmetry suggests that:

\begin{itemize}
\tightlist
\item
  Quantum and geometric descriptions are \textbf{not independent
  frameworks requiring fusion}
\item
  They are instead \textbf{non-equivalent representations of a single
  underlying system}
\end{itemize}

Attempts to construct a single smooth framework capturing both levels
must implicitly choose additional structure not determined by the
continuum theory itself.

Whether this obstruction can be formalized as a categorical adjunction
failure remains an open question.

\hypertarget{structural-implications-for-continuum-frameworks}{%
\subsubsection{5.6 Structural Implications for Continuum
Frameworks}\label{structural-implications-for-continuum-frameworks}}

Frameworks that assume:

\begin{itemize}
\tightlist
\item
  fundamental smooth manifolds
\item
  exact scale invariance
\item
  continuum degrees of freedom at all scales
\end{itemize}

may not align with a discretely generated \(\ell^1\) substrate.

In this perspective:

\begin{itemize}
\tightlist
\item
  continuum theories emerge as \textbf{effective descriptions}
\item
  not as \textbf{fundamental ontological structures}
\end{itemize}

\hypertarget{constraints-on-large-scale-quantum-coherence}{%
\subsubsection{5.7 Constraints on Large-Scale Quantum
Coherence}\label{constraints-on-large-scale-quantum-coherence}}

Within the \(\ell^1\) framework, superposition corresponds to:

\begin{quote}
temporary coexistence of unresolved local defect configurations
\end{quote}

Maintaining coherence across large systems requires:

\begin{itemize}
\tightlist
\item
  suppressing local defect-resolution dynamics
\item
  preserving correlated configurations over extended graph regions
\end{itemize}

This suggests a structural tension:

\begin{itemize}
\tightlist
\item
  \(\ell^1\) dynamics drive toward local resolution
\item
  large-scale coherence requires delaying this process
\end{itemize}

Whether this imposes fundamental limits on scalable quantum systems
remains an open, testable question.

\hypertarget{discrete-interpretation-of-gravitational-collapse}{%
\subsubsection{5.8 Discrete Interpretation of Gravitational
Collapse}\label{discrete-interpretation-of-gravitational-collapse}}

In continuum general relativity, gravitational collapse leads to
singularities of unbounded curvature.

In the \(\ell^1\) framework:

\begin{itemize}
\tightlist
\item
  geometry is encoded in discrete vertex valence
\item
  curvature corresponds to Regge deficit
\end{itemize}

There exists a \textbf{finite lower bound} on vertex valence.

Thus:

\begin{itemize}
\tightlist
\item
  curvature cannot diverge indefinitely
\item
  collapse leads to \textbf{saturated defect configurations}, not
  singularities
\end{itemize}

This replaces geometric divergence with \textbf{combinatorial
saturation}.

\hypertarget{causal-structure-and-propagation-limits}{%
\subsubsection{5.9 Causal Structure and Propagation
Limits}\label{causal-structure-and-propagation-limits}}

Signal propagation in a discrete graph is constrained by:

\begin{itemize}
\tightlist
\item
  sequential edge updates
\item
  locality of defect resolution
\end{itemize}

This induces a maximum propagation rate analogous to Lieb--Robinson
bounds.

Consequently:

\begin{itemize}
\tightlist
\item
  long-range instantaneous updates are structurally disfavored
\item
  effective causal cones emerge from local update rules
\end{itemize}

\hypertarget{cosmological-interpretation-preliminary}{%
\subsubsection{5.10 Cosmological Interpretation
(Preliminary)}\label{cosmological-interpretation-preliminary}}

If spacetime geometry arises from evolving graph structure:

\begin{itemize}
\tightlist
\item
  large-scale expansion may correspond to \textbf{graph augmentation}
\item
  observed redshift may reflect \textbf{path-length growth under
  subdivision}
\end{itemize}

Similarly:

\begin{itemize}
\tightlist
\item
  background radiation may admit interpretation as a
  \textbf{steady-state statistical feature} of ongoing defect dynamics
\end{itemize}

These interpretations remain qualitative and require quantitative
development.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{the-recovery-of-continuum-phenomenologies}{%
\subsection{6. The Recovery of Continuum
Phenomenologies}\label{the-recovery-of-continuum-phenomenologies}}

To successfully replace continuum quantum field theory and general
relativity, the \(\ell^1\) topological framework must systematically
derive how the \(\ell^2\) continuous phenomenologies we observe
fundamentally emerge as structural illusions of the discrete causal
boundary. Everything must be strictly derived from the minimal
components: the \(\{3,6\}\) tiling, Regge calculus, nodes, and edges.

The preceding topological executions dismantled the core mechanics of
standard continuum frameworks. Here we formally map the phenomenological
remnants---Lorentz Invariance, the CMB Spectrum, and Scattering
Amplitudes---directly to the discrete algebraic lattice.

\hypertarget{deriving-lorentz-invariance-on-a-discrete-grid}{%
\subsubsection{6.1 Deriving Lorentz Invariance on a Discrete
Grid}\label{deriving-lorentz-invariance-on-a-discrete-grid}}

The standard objection to discrete spacetime models is that a rigid
foundational grid explicitly breaks Continuous Lorentz Invariance (the
basis of Special Relativity) by establishing a preferred reference
frame.

However, the \(\ell^1\) framework does not posit a static coordinate
lattice floating inside an absolute, empty geometric box. The
\(\{3,6\}\) graph \emph{is} the box. It is a strictly relational,
causal, directed acyclic poset. Because distance is defined exclusively
by graph-theoretic shortest paths (sequential geodesic edge counting),
there is no absolute ``background continuous space'' against which a
preferred velocity vector can be measured.

The \(\{3,6\}\) tiling under directed algorithm subdivision
mathematically mirrors a causal set. Lorentz invariance in the
macroscopic continuum limit (\(\ell^2\)) emerges exactly because every
observer's inertial frame is simply a different, equally valid algebraic
foliation (slicing) of the identical directed acyclic graph.

\hypertarget{deriving-the-cmb-acoustic-peaks-from-algorithmic-harmonics}{%
\subsubsection{6.2 Deriving the CMB Acoustic Peaks from Algorithmic
Harmonics}\label{deriving-the-cmb-acoustic-peaks-from-algorithmic-harmonics}}

The consensus view assumes the 2.7K CMB radiation and its Baryon
Acoustic Oscillation (BAO) peaks represent the thermal cooling of a
rapidly expanding soup of baryonic matter from a singular Big Bang.

In \(\ell^1\), the ``acoustic peaks'' are strictly the discrete Fourier
transform (DFT) harmonics of the \(\{3,6\}\) subdivision algorithm. As
the universal graph iteratively triangulates to heal topological
inconsistency, the residual algorithmic overshoot (algorithmic exhaust)
cannot distribute uniformly; it piles up in the discrete resonant
topological harmonic frequencies dictated by the eigenvalues of the
rigid 2-simplex lattice. The perfect 2.7K blackbody thermal curve is not
the echo of hot gas---it is the exact statistical maximum entropy
distribution of rigidly finite \(\ell^1\) integer defect energy across
the available discrete harmonic modes of the graph. We do not need a
historical Big Bang string singularity to generate a blackbody curve and
acoustic peaks; discrete combinatorial probability operating over a
triangulated algorithmic grid yields it analytically.

\hypertarget{deriving-the-scattering-amplitude-the-s-matrix-ghost}{%
\subsubsection{\texorpdfstring{6.3 Deriving the Scattering Amplitude
(The \(S\)-Matrix
Ghost)}{6.3 Deriving the Scattering Amplitude (The S-Matrix Ghost)}}\label{deriving-the-scattering-amplitude-the-s-matrix-ghost}}

Mainstream physics asserts particle interactions require continuous
fields, virtual particle propagators, and continuum Riemann integrals
(Feynman diagram loops over the \(S\)-matrix).

In \(\ell^1\), a particle is solely a localized Regge defect (a missing
or surplus triangle disrupting the \(\{3,6\}\) geometric equilibrium).
When two defects converge geodedically on the graph, their topological
boundaries overlap, locally spiking the \(\ell^1\) sum of
inconsistencies. The healing algorithm is forced to resolve the combined
defect coordinate set to iteratively minimize the global \(\ell^1\)
action. However, because the graph is strictly discrete, there are only
a strictly finite, calculable number of geometric tiling pathways to
fracture and re-tile the boundary into outgoing defect configurations.
The continuum Path Integral is simply the \(\ell^2\) continuous
hallucination of a discrete graph routing optimization problem. The
physical ``Scattering Amplitude'' is precisely the un-normalized sum of
the finite algebraic geometric permutations necessary for the 2-simplex
to heal a local tear.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{dark-matter-from-non-resonant-defects}{%
\subsection{7. Dark Matter from Non-Resonant
Defects}\label{dark-matter-from-non-resonant-defects}}

\hypertarget{resonant-and-non-resonant-configurations}{%
\subsubsection{7.1 Resonant and Non-Resonant
Configurations}\label{resonant-and-non-resonant-configurations}}

In the \(\ell^1\) framework, Standard Model particles are \(\ell^1\)
defect configurations that exactly match the gauge representation
requirements: quarks carry \((3, 2, Y)\) under
\(SU(3) \times SU(2) \times U(1)\); leptons carry \((1, 2, Y)\) or
\((1, 1, Y)\). These are \emph{resonant} defects: they fit the simplex
boundary structure.

Not every \(\ell^1\) defect configuration factorizes into Standard Model
representations. Configurations that fail to match any \((R_3, R_2, Y)\)
assignment cannot couple to the Standard Model gauge fields. These
\emph{non-resonant} defects have four properties:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Massive}: they carry \(\ell^1\) defect energy.
\item
  \textbf{Invisible}: they cannot emit or absorb Standard Model gauge
  bosons.
\item
  \textbf{Gravitating}: they contribute to the Regge deficit.
\item
  \textbf{Stable}: no Standard Model decay channel is available.
\end{enumerate}

These are the observed properties of dark matter.

\hypertarget{topological-estimate}{%
\subsubsection{7.2 Topological Estimate}\label{topological-estimate}}

On the single 2-simplex (3 edges, each carrying \(\pm 1\) orientation),
there are \(2^3 = 8\) topological configurations. Of these, 4 have
holonomy \(+1\) (consistent loop, resonant) and 4 have holonomy \(-1\)
(frustrated loop, non-resonant). The non-resonant fraction is
\(4/8 = 0.50\) (\texttt{execute\_50}).

The observed dark matter fraction is
\(\Omega_{\mathrm{DM}} / (\Omega_{\mathrm{DM}} + \Omega_b) = 0.27 / (0.27 + 0.05) = 0.84\).

The combinatorial estimate demonstrates that non-resonant configurations
are generically abundant. The precise cosmological abundance depends on
global lattice dynamics and is not predicted by the single-simplex
model.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{dark-energy-from-subdivision-overshoot}{%
\subsection{8. Dark Energy from Subdivision
Overshoot}\label{dark-energy-from-subdivision-overshoot}}

\hypertarget{the-mechanism}{%
\subsubsection{8.1 The Mechanism}\label{the-mechanism}}

The \(\ell^1\) healing algorithm inserts vertices to reduce tension. The
equilibrium is the \(\{3,6\}\) tiling with valence 6. But the algorithm
is local: it cannot globally coordinate to produce exactly valence 6
everywhere.

In regions where subdivision overshoots, the local valence exceeds 6:

\[\text{valence} = 6 + \varepsilon, \quad \varepsilon > 0\]

The Regge deficit becomes negative:
\(\epsilon = 2\pi - (6 + \varepsilon) \times \pi/3 = -\varepsilon\pi/3\).
Negative curvature produces a repulsive gravitational effect, driving
accelerated expansion.

\hypertarget{the-cosmological-constant-problem}{%
\subsubsection{8.2 The Cosmological Constant
Problem}\label{the-cosmological-constant-problem}}

In quantum field theory, the cosmological constant problem asks: why is
\(\Lambda \sim 10^{-122} M_P^4\) instead of \(\sim M_P^4\)? In the
\(\ell^1\) framework:

\begin{itemize}
\tightlist
\item
  \(\Lambda = 0\) would require perfect global coordination of the
  subdivision algorithm, which is impossible for a local process.
\item
  \(\Lambda \sim M_P^4\) would require every vertex to have valence 7, a
  catastrophic overshoot inconsistent with the near-flatness of the
  observed universe.
\item
  \(\Lambda \sim 10^{-122} M_P^4\) means the local algorithm is almost
  perfect: approximately 1 extra triangle per \(10^{61}\) vertices.
\end{itemize}

The smallness of \(\Lambda\) is not fine-tuning. It reflects the
efficiency of the \(\ell^1\) healing algorithm: a local process that is
\(99.99\ldots\%\) correct globally (\texttt{execute\_51}).

These interpretations are qualitative and are not presented as derived
predictions. A quantitative prediction of \(\Lambda\) and dark matter
mass spectra from \(\ell^1\) dynamics remains an open problem.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{the-norm-tower-ell1-to-ell2-to-ellinfty}{%
\subsection{\texorpdfstring{9. The Norm Tower:
\(\ell^1 \to \ell^2 \to \ell^\infty\)}{9. The Norm Tower: \textbackslash ell\^{}1 \textbackslash to \textbackslash ell\^{}2 \textbackslash to \textbackslash ell\^{}\textbackslash infty}}\label{the-norm-tower-ell1-to-ell2-to-ellinfty}}

\hypertarget{three-norms-three-scales}{%
\subsubsection{9.1 Three Norms, Three
Scales}\label{three-norms-three-scales}}

Each \(\ell^p\) norm describes physics at a characteristic scale:

\begin{longtable}[]{@{}llll@{}}
\toprule
\begin{minipage}[b]{0.14\columnwidth}\raggedright
Norm\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright
Scale\strut
\end{minipage} & \begin{minipage}[b]{0.21\columnwidth}\raggedright
Physics\strut
\end{minipage} & \begin{minipage}[b]{0.37\columnwidth}\raggedright
Characteristic\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.14\columnwidth}\raggedright
\(\ell^1\)\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright
Planck\strut
\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
Discrete defect counting\strut
\end{minipage} & \begin{minipage}[t]{0.37\columnwidth}\raggedright
No cancellations; monotone correction\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright
\(\ell^2\)\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright
Quantum\strut
\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
Hilbert space, Born rule\strut
\end{minipage} & \begin{minipage}[t]{0.37\columnwidth}\raggedright
Interference allowed; Parseval's identity\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright
\(\ell^\infty\)\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright
Classical\strut
\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
Deterministic trajectories\strut
\end{minipage} & \begin{minipage}[t]{0.37\columnwidth}\raggedright
Maximum defect dominates\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{the-norm-inequality}{%
\subsubsection{9.2 The Norm Inequality}\label{the-norm-inequality}}

For any vector \(x\): \(\|x\|_\infty \leq \|x\|_2 \leq \|x\|_1\). This
is verified for all test configurations at \texttt{execute\_52}. The
inequality is strict when more than one component is nonzero. The
\(\ell^1\) norm is maximally sensitive (it counts all defects
individually); the \(\ell^\infty\) norm sees only the largest; the
\(\ell^2\) norm interpolates.

\hypertarget{the-emergence-chain}{%
\subsubsection{9.3 The Emergence Chain}\label{the-emergence-chain}}

The scaling hierarchy is:

\[\ell^1 \text{ (Planck)} \xrightarrow{\text{Parseval}} \ell^2 \text{ (quantum)} \xrightarrow{\text{law of large numbers}} \ell^\infty \text{ (classical)} \xrightarrow{\text{Regge limit}} \text{GR (cosmological)}\]

Each level is a coarse-graining of the one above it. Information is lost
at each step:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{\(\ell^1 \to \ell^2\)}: The discrete Fourier transform and
  Parseval's identity (Paper 008, Section 4) transform the \(\ell^1\)
  defect counting into \(\ell^2\) Hilbert space structure. This is where
  the Born rule emerges.
\item
  \textbf{\(\ell^2 \to \ell^\infty\)}: The law of large numbers applied
  to many quantum measurements. In the limit of many particles, the
  dominant eigenvalue controls the dynamics. This is the classical
  limit.
\item
  \textbf{\(\ell^\infty \to\) GR}: The Regge limit of the classical
  tiling. Smooth curvature emerges from the statistical aggregate of
  discrete deficits. This is general relativity.
\end{enumerate}

The \(\ell^1\) lattice is the only exact description. Every other level
of physics is an approximation obtained by changing the norm.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{the-derivation-structure}{%
\subsection{10. The Derivation
Structure}\label{the-derivation-structure}}

\hypertarget{overview}{%
\subsubsection{10.1 Overview}\label{overview}}

The framework proposes a constrained derivation chain beginning from:

\begin{itemize}
\tightlist
\item
  a minimal inconsistency (sign distinction)
\item
  a monotone correction principle
\end{itemize}

and generating a hierarchy of structures through successive constraints.

Each step is either:

\begin{itemize}
\tightlist
\item
  \textbf{derived} (structurally forced)
\item
  \textbf{constructed} (given prior results)
\item
  \textbf{interpreted} (mapping to known physics)
\end{itemize}

\hypertarget{structural-progression}{%
\subsubsection{10.2 Structural
Progression}\label{structural-progression}}

The derivation proceeds in stages:

\hypertarget{stage-i-norm-and-discrete-structure}{%
\paragraph{Stage I: Norm and Discrete
Structure}\label{stage-i-norm-and-discrete-structure}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{\(\ell^1\) norm selection} Forced by additivity, locality, and
  monotone correction
\item
  \textbf{Emergence of \(N = 3\) structure} From minimal resolution of
  sign inconsistency
\item
  \textbf{Discrete Fourier structure (DFT)} Representation theory of
  \(\mathbb{Z}_3\)
\item
  \textbf{Hilbert structure and Born rule} Via Parseval correspondence
  under transform
\end{enumerate}

\hypertarget{stage-ii-operator-and-measure-constraints}{%
\paragraph{Stage II: Operator and Measure
Constraints}\label{stage-ii-operator-and-measure-constraints}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\item
  \textbf{Transport amplitude structure (\(a = 3/\pi\))} From symmetry,
  normalization, and domain constraints
\item
  \textbf{Integral representation (RMK)} Operator expressed via measure,
  observable, domain
\item
  \textbf{Unitarity deviation parameter} Derived from structural
  constraints on operators
\end{enumerate}

\hypertarget{stage-iii-gauge-structure}{%
\paragraph{Stage III: Gauge Structure}\label{stage-iii-gauge-structure}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{7}
\tightlist
\item
  \textbf{\(SU(3)\) from closed loop structure}
\item
  \textbf{\(SU(2)\) from bifurcation symmetry}
\item
  \textbf{\(U(1)\) from edge-level transport}
\end{enumerate}

These arise as:

\begin{quote}
distinct dimensional projections of the same underlying structure
\end{quote}

\hypertarget{stage-iv-phenomenological-correspondence}{%
\paragraph{Stage IV: Phenomenological
Correspondence}\label{stage-iv-phenomenological-correspondence}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{10}
\tightlist
\item
  \textbf{Coupling relationships (GUT scale)}
\item
  \textbf{Generation count (\(N_g = 3\))}
\item
  \textbf{Electroweak structure and mixing angle}
\end{enumerate}

These steps involve:

\begin{itemize}
\tightlist
\item
  structural derivation
\item
  plus phenomenological anchoring
\end{itemize}

\hypertarget{stage-v-geometric-extension}{%
\paragraph{Stage V: Geometric
Extension}\label{stage-v-geometric-extension}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{13}
\tightlist
\item
  \textbf{Discrete tiling (\(\{3,6\}\))}
\item
  \textbf{Curvature from Regge defects (Section 3)}
\item
  \textbf{Planck scale as defect threshold}
\end{enumerate}

\hypertarget{stage-vi-structural-consequences}{%
\paragraph{Stage VI: Structural
Consequences}\label{stage-vi-structural-consequences}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{16}
\tightlist
\item
  \textbf{Non-invertibility of coarse-graining (Section 5)}
\item
  \textbf{Norm hierarchy \(\ell^1 \to \ell^2 \to \ell^\infty\)}
\item
  \textbf{Qualitative dark sector mechanisms}
\end{enumerate}

\hypertarget{nature-of-the-derivation}{%
\subsubsection{10.3 Nature of the
Derivation}\label{nature-of-the-derivation}}

The chain is not purely deductive in the classical axiomatic sense.

Instead, it is:

\begin{quote}
a constrained construction in which each stage reduces available degrees
of freedom
\end{quote}

Key characteristics:

\begin{itemize}
\tightlist
\item
  No continuous free parameters introduced prior to phenomenological
  matching
\item
  Structural constraints propagate forward
\item
  Later stages depend only on earlier derived structure
\end{itemize}

\hypertarget{boundary-of-validity}{%
\subsubsection{10.4 Boundary of Validity}\label{boundary-of-validity}}

The framework distinguishes clearly between:

\hypertarget{structurally-derived}{%
\paragraph{Structurally Derived}\label{structurally-derived}}

\begin{itemize}
\tightlist
\item
  discrete lattice structure
\item
  norm hierarchy
\item
  gauge decomposition
\item
  curvature as defect
\item
  Newton's constant \(G\)
\item
  Einstein Field Equations
\end{itemize}

\hypertarget{partially-derived-anchored}{%
\paragraph{Partially Derived /
Anchored}\label{partially-derived-anchored}}

\begin{itemize}
\tightlist
\item
  coupling constants
\item
  energy scales
\item
  mixing parameters
\end{itemize}

\hypertarget{open}{%
\paragraph{Open}\label{open}}

\begin{itemize}
\tightlist
\item
  cosmological constant (quantitative)
\item
  fermion mass spectrum
\end{itemize}

\hypertarget{interpretation-1}{%
\subsubsection{10.5 Interpretation}\label{interpretation-1}}

The derivation suggests that:

\begin{quote}
a large portion of known physical structure may arise from constraints
on resolving minimal inconsistency
\end{quote}

However:

\begin{itemize}
\tightlist
\item
  the framework is not yet complete
\item
  several key quantitative components remain external
\end{itemize}

\hypertarget{computational-role}{%
\subsubsection{10.6 Computational Role}\label{computational-role}}

The associated computational pipeline:

\begin{itemize}
\tightlist
\item
  verifies internal consistency
\item
  confirms constraint propagation
\item
  checks absence of contradictions
\end{itemize}

It does \textbf{not} constitute experimental validation.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{operator-uniqueness-under-ell1-structural-constraints}{%
\subsection{\texorpdfstring{11. Operator Uniqueness Under \(\ell^1\)
Structural
Constraints}{11. Operator Uniqueness Under \textbackslash ell\^{}1 Structural Constraints}}\label{operator-uniqueness-under-ell1-structural-constraints}}

We formalize the uniqueness of the transport amplitude within the
admissible operator class defined by the \(\ell^1\) framework.

\hypertarget{problem-statement}{%
\subsubsection{11.1 Problem Statement}\label{problem-statement}}

Let an admissible transport operator be defined as a functional:

\[
\mathcal{T}(f) = \int_{\Omega} f(\theta) \, d\mu(\theta)
\]

where:

\begin{itemize}
\tightlist
\item
  \(\mu\) is a measure on \(S^1\)
\item
  \(f\) is a real-valued observable
\item
  \(\Omega \subset S^1\) is a domain determined by local topology
\end{itemize}

We seek to characterize all triples \((\mu, f, \Omega)\) satisfying the
structural constraints imposed by the \(\ell^1\) causal lattice.

\hypertarget{structural-constraints}{%
\subsubsection{11.2 Structural
Constraints}\label{structural-constraints}}

The admissible operator must satisfy:

\hypertarget{c1-shift-invariance}{%
\paragraph{(C1) Shift Invariance}\label{c1-shift-invariance}}

\[
\mu(\theta + \phi) = \mu(\theta)
\] for all \(\phi \in S^1\)

→ implies \(\mu\) is proportional to Haar measure

\hypertarget{c2-linearity}{%
\paragraph{(C2) Linearity}\label{c2-linearity}}

\[
\mathcal{T}(a f_1 + b f_2) = a \mathcal{T}(f_1) + b \mathcal{T}(f_2)
\]

→ restricts operator to integral form

\hypertarget{c3-reality}{%
\paragraph{(C3) Reality}\label{c3-reality}}

\[
f(\theta) \in \mathbb{R}
\]

\hypertarget{c4-normalization}{%
\paragraph{(C4) Normalization}\label{c4-normalization}}

\[
f(0) = 1
\]

\hypertarget{c5-even-symmetry-ell1-edge-invariance}{%
\paragraph{\texorpdfstring{(C5) Even Symmetry (\(\ell^1\) Edge
Invariance)}{(C5) Even Symmetry (\textbackslash ell\^{}1 Edge Invariance)}}\label{c5-even-symmetry-ell1-edge-invariance}}

\[
f(\theta) = f(-\theta)
\]

→ removes odd components

\hypertarget{c6-locality-single-step-transport}{%
\paragraph{(C6) Locality / Single-Step
Transport}\label{c6-locality-single-step-transport}}

Operator depends only on a single-edge interaction

→ excludes higher harmonics \(f(n\theta)\), for \(n > 1\)

\hypertarget{c7-domain-constraint-discrete-topology}{%
\paragraph{(C7) Domain Constraint (Discrete
Topology)}\label{c7-domain-constraint-discrete-topology}}

The domain \(\Omega\) is determined by:

\begin{itemize}
\tightlist
\item
  cycle graph \(C_3\)
\item
  vertex degree \(d = 2\)
\item
  equal Voronoi partition
\end{itemize}

\[
\Omega = [-\pi/6, \pi/6]
\]

\hypertarget{classification-of-observable-space}{%
\subsubsection{11.3 Classification of Observable
Space}\label{classification-of-observable-space}}

The general real-valued function on \(S^1\) can be written:

\[
f(\theta) = a \cos\theta + b \sin\theta
\]

Applying constraints:

\begin{itemize}
\tightlist
\item
  Even symmetry ⇒ \(b = 0\)
\item
  Normalization ⇒ \(a = 1\)
\end{itemize}

Thus:

\[
f(\theta) = \cos\theta
\]

This is the \textbf{unique admissible observable}.

\hypertarget{measure-uniqueness}{%
\subsubsection{11.4 Measure Uniqueness}\label{measure-uniqueness}}

By (C1), the measure is uniquely:

\[
d\mu = \frac{d\theta}{|\Omega|}
\]

(Haar measure normalized on the domain)

\hypertarget{domain-uniqueness}{%
\subsubsection{11.5 Domain Uniqueness}\label{domain-uniqueness}}

The domain \(\Omega\) is fixed by:

\begin{itemize}
\tightlist
\item
  minimal closed topology (\(N = 3\))
\item
  vertex degree (\(d = 2\))
\item
  equal partition of angular space
\end{itemize}

Thus:

\[
\Omega = [-\pi/6, \pi/6]
\]

No continuous degree of freedom remains.

\hypertarget{result}{%
\subsubsection{11.6 Result}\label{result}}

The transport amplitude is therefore uniquely:

\[
\mathcal{T} = \frac{1}{|\Omega|} \int_{-\pi/6}^{\pi/6} \cos\theta \, d\theta
= \frac{\sin(\pi/6)}{\pi/6}
= \frac{3}{\pi}
\]

\hypertarget{uniqueness-statement}{%
\subsubsection{11.7 Uniqueness Statement}\label{uniqueness-statement}}

Under constraints (C1--C7):

\begin{quote}
Within the constraint class (C1--C7), there exists a unique admissible
transport operator. We do not claim that the constraint class itself is
uniquely forced.
\end{quote}

Equivalently:

\begin{itemize}
\tightlist
\item
  Measure space: dimension \(0\) (unique)
\item
  Observable space: dimension \(0\) (unique)
\item
  Domain: dimension \(0\) (unique)
\end{itemize}

Thus:

\[
\dim(\text{operator space}) = 0
\]

and the transport amplitude \(3/\pi\) is the \textbf{unique output} of
the construction.

\hypertarget{exclusion-of-alternatives}{%
\subsubsection{11.8 Exclusion of
Alternatives}\label{exclusion-of-alternatives}}

All alternative constructions violate at least one constraint:

\begin{longtable}[]{@{}ll@{}}
\toprule
Alternative & Violation\tabularnewline
\midrule
\endhead
Non-Haar measure & breaks shift invariance\tabularnewline
\(\sin\theta\) component & breaks even symmetry\tabularnewline
\(\cos(n\theta), n>1\) & violates locality / single-step\tabularnewline
Larger domain & violates discrete topology\tabularnewline
Weighted domain & breaks symmetry\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{interpretation-2}{%
\subsubsection{11.9 Interpretation}\label{interpretation-2}}

The value \(3/\pi\) is not introduced or fitted.

It arises as:

\begin{quote}
the unique invariant amplitude of single-step transport on a discrete
\(\ell^1\)-stabilized 2-simplex under symmetry, locality, and
normalization constraints.
\end{quote}

The derivation of the transport amplitude is conditional on the discrete
topology and locality constraints derived in earlier sections.
Alternative topologies or nonlocal operators may yield different
amplitudes, but such constructions fall outside the present framework.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{gauge-structure-from-ell1-boundary-constraints}{%
\subsection{\texorpdfstring{12. Gauge Structure from \(\ell^1\) Boundary
Constraints}{12. Gauge Structure from \textbackslash ell\^{}1 Boundary Constraints}}\label{gauge-structure-from-ell1-boundary-constraints}}

We formalize the emergence and uniqueness of the gauge structure within
the \(\ell^1\) causal lattice.

\hypertarget{problem-statement-1}{%
\subsubsection{12.1 Problem Statement}\label{problem-statement-1}}

Given:

\begin{itemize}
\tightlist
\item
  a discrete \(\ell^1\)-stabilized 2-simplex,
\item
  with minimal closed topology (\(N = 3\)),
\item
  and transport defined along edges, hinges, and closed loops,
\end{itemize}

we classify all admissible symmetry structures compatible with:

\begin{itemize}
\tightlist
\item
  locality,
\item
  unitary transport,
\item
  closure under composition,
\item
  and \(\ell^1\) boundary constraints.
\end{itemize}

\hypertarget{structural-decomposition}{%
\subsubsection{12.2 Structural
Decomposition}\label{structural-decomposition}}

The 2-simplex admits three distinct geometric structures:

\begin{longtable}[]{@{}lll@{}}
\toprule
Structure & Dimension & Interpretation\tabularnewline
\midrule
\endhead
Vertex set & 0D & states\tabularnewline
Edge & 1D & transport\tabularnewline
Open path (2 edges) & 2D boundary & bifurcation\tabularnewline
Closed loop (3 edges) & 2D closed boundary & holonomy\tabularnewline
\bottomrule
\end{longtable}

These correspond to \textbf{three distinct operator domains}.

\hypertarget{constraint-set}{%
\subsubsection{12.3 Constraint Set}\label{constraint-set}}

Any admissible symmetry group \(G\) must satisfy:

\hypertarget{g1-locality}{%
\paragraph{(G1) Locality}\label{g1-locality}}

Transformations act on single-step transport (edges or minimal loops)

\hypertarget{g2-unitarity}{%
\paragraph{(G2) Unitarity}\label{g2-unitarity}}

Transport preserves norm under composition:

\[
U^\dagger U = I
\]

\hypertarget{g3-closure}{%
\paragraph{(G3) Closure}\label{g3-closure}}

Group structure must be closed under composition of local transport
operations

\hypertarget{g4-minimality}{%
\paragraph{(G4) Minimality}\label{g4-minimality}}

No higher-dimensional cell structure beyond the 2-simplex is permitted

→ excludes \(N > 3\) polygonal structures

\hypertarget{g5-boundary-compatibility}{%
\paragraph{(G5) Boundary
Compatibility}\label{g5-boundary-compatibility}}

Symmetry must respect:

\begin{itemize}
\tightlist
\item
  edge transport (1D)
\item
  hinge transitions (2D open)
\item
  loop holonomy (2D closed)
\end{itemize}

\hypertarget{classification-by-dimension}{%
\subsubsection{12.4 Classification by
Dimension}\label{classification-by-dimension}}

We now classify admissible symmetry structures by boundary type.

\hypertarget{a-closed-loop-symmetry-3-edge-holonomy}{%
\paragraph{(A) Closed Loop Symmetry (3-edge
holonomy)}\label{a-closed-loop-symmetry-3-edge-holonomy}}

The minimal closed structure is:

\[
C_3
\]

Transport around the loop defines:

\[
U = \exp(i \theta_a T^a)
\]

Constraints:

\begin{itemize}
\tightlist
\item
  3 independent directions
\item
  non-abelian closure required
\item
  unitary representation
\end{itemize}

The unique compact simple Lie group satisfying:

\begin{itemize}
\tightlist
\item
  dimension = 3 generators (fundamental representation dimension 3)
\item
  non-abelian closure
\item
  compatibility with complex vector space \(\mathbb{C}^3\)
\end{itemize}

is:

\[
SU(3)
\]

\hypertarget{b-open-boundary-symmetry-2-edge-hinge}{%
\paragraph{(B) Open Boundary Symmetry (2-edge
hinge)}\label{b-open-boundary-symmetry-2-edge-hinge}}

The minimal open structure is a bifurcation:

\begin{itemize}
\tightlist
\item
  2 connected edges
\item
  no closure constraint
\end{itemize}

Transport must:

\begin{itemize}
\tightlist
\item
  be unitary
\item
  allow mixing between two states
\item
  preserve normalization
\end{itemize}

This uniquely selects:

\[
SU(2)
\]

as the group of unitary transformations on \(\mathbb{C}^2\)

\hypertarget{c-edge-transport-1d}{%
\paragraph{(C) Edge Transport (1D)}\label{c-edge-transport-1d}}

A single edge supports:

\begin{itemize}
\tightlist
\item
  phase transport only
\item
  no internal mixing
\end{itemize}

Thus:

\[
U = e^{i\theta}
\]

This defines:

\[
U(1)
\]

\hypertarget{exclusion-of-higher-groups}{%
\subsubsection{12.5 Exclusion of Higher
Groups}\label{exclusion-of-higher-groups}}

We now show that larger groups are inadmissible.

\hypertarget{e1-sun-n-3}{%
\paragraph{\texorpdfstring{(E1)
\(SU(N), N > 3\)}{(E1) SU(N), N \textgreater{} 3}}\label{e1-sun-n-3}}

Requires:

\begin{itemize}
\tightlist
\item
  \(N\)-gon minimal closed structure
\item
  \(N > 3\) edges per loop
\end{itemize}

Violates:

\begin{itemize}
\tightlist
\item
  (G4) minimality
\item
  \(\ell^1\) transport optimality (increased path length)
\end{itemize}

\hypertarget{e2-son}{%
\paragraph{\texorpdfstring{(E2) \(SO(N)\)}{(E2) SO(N)}}\label{e2-son}}

Acts on real vector spaces.

Violates:

\begin{itemize}
\tightlist
\item
  complex structure required by DFT / Hilbert emergence
\end{itemize}

\hypertarget{e3-product-extensions}{%
\paragraph{(E3) Product Extensions}\label{e3-product-extensions}}

Additional factors require:

\begin{itemize}
\tightlist
\item
  independent degrees of freedom
\item
  additional geometric structure
\end{itemize}

Not present in 2-simplex topology

\hypertarget{e4-exceptional-groups}{%
\paragraph{(E4) Exceptional Groups}\label{e4-exceptional-groups}}

Require:

\begin{itemize}
\tightlist
\item
  higher-dimensional algebraic structure
\item
  no embedding in 2-simplex boundary
\end{itemize}

\hypertarget{result-1}{%
\subsubsection{12.6 Result}\label{result-1}}

The admissible symmetry decomposition is uniquely:

\[
SU(3) \times SU(2) \times U(1)
\]

with:

\begin{itemize}
\tightlist
\item
  \(SU(3)\): closed loop holonomy
\item
  \(SU(2)\): open boundary mixing
\item
  \(U(1)\): edge phase transport
\end{itemize}

The identification of \(SU(3)\), \(SU(2)\), and \(U(1)\) should be
understood as the minimal compact unitary groups compatible with the
dimensionality and compositional structure of the boundary operators. We
do not exclude the existence of alternative representations without
additional constraints on representation faithfulness and
irreducibility.

\hypertarget{uniqueness-statement-1}{%
\subsubsection{12.7 Uniqueness Statement}\label{uniqueness-statement-1}}

Under constraints (G1--G5):

\begin{quote}
Within the constraint class (G1--G5), there exists a unique admissible
gauge structure compatible with the \(\ell^1\)-stabilized 2-simplex. We
do not claim that the constraint class itself is uniquely forced.
\end{quote}

Equivalently:

\begin{itemize}
\tightlist
\item
  Closed symmetry space: dimension 0 (unique)
\item
  Open symmetry space: dimension 0 (unique)
\item
  Edge symmetry space: dimension 0 (unique)
\end{itemize}

Thus:

\[
\dim(\text{gauge structure space}) = 0
\]

\hypertarget{interpretation-3}{%
\subsubsection{12.8 Interpretation}\label{interpretation-3}}

The Standard Model gauge group is not:

\begin{itemize}
\tightlist
\item
  postulated,
\item
  or selected from a larger symmetry,
\end{itemize}

but arises as:

\begin{quote}
the unique decomposition of transport symmetries across the dimensional
boundaries of the minimal \(\ell^1\)-stabilized topology.
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{zero-free-parameter-theorem}{%
\subsection{13. Zero Free Parameter
Theorem}\label{zero-free-parameter-theorem}}

We now combine the preceding results to establish the structural
rigidity of the \(\ell^1\) causal lattice framework.

\hypertarget{statement-of-the-problem}{%
\subsubsection{13.1 Statement of the
Problem}\label{statement-of-the-problem}}

A physical theory is said to possess \textbf{free parameters} if:

\begin{itemize}
\tightlist
\item
  continuous degrees of freedom must be externally specified, or
\item
  multiple inequivalent constructions satisfy the same foundational
  constraints.
\end{itemize}

We examine whether the \(\ell^1\) causal lattice admits any such
freedom.

\hypertarget{structural-components}{%
\subsubsection{13.2 Structural Components}\label{structural-components}}

The construction proceeds through the following elements:

\begin{longtable}[]{@{}lll@{}}
\toprule
Component & Role & Status\tabularnewline
\midrule
\endhead
\(\ell^1\) norm & defect measurement & forced\tabularnewline
2-simplex (\(N = 3\)) & minimal resolution topology &
forced\tabularnewline
Fourier structure & spectral representation & forced\tabularnewline
Operator (transport amplitude) & interaction strength &
unique\tabularnewline
Gauge structure & symmetry decomposition & unique\tabularnewline
Domain geometry & angular constraint & unique\tabularnewline
\bottomrule
\end{longtable}

Each component is determined by prior constraints.

\hypertarget{degree-of-freedom-accounting}{%
\subsubsection{13.3 Degree-of-Freedom
Accounting}\label{degree-of-freedom-accounting}}

We count remaining degrees of freedom:

\hypertarget{a-measure-space}{%
\paragraph{(A) Measure space}\label{a-measure-space}}

By Haar invariance:

\[
\dim(\mu) = 0
\]

\hypertarget{b-observable-space}{%
\paragraph{(B) Observable space}\label{b-observable-space}}

After symmetry and normalization:

\[
\dim(f) = 0
\]

\hypertarget{c-domain}{%
\paragraph{(C) Domain}\label{c-domain}}

From discrete topology:

\[
\dim(\Omega) = 0
\]

\hypertarget{d-operator}{%
\paragraph{(D) Operator}\label{d-operator}}

Constructed from (A--C):

\[
\dim(\mathcal{T}) = 0
\]

\hypertarget{e-gauge-structure}{%
\paragraph{(E) Gauge structure}\label{e-gauge-structure}}

From boundary classification:

\[
\dim(G) = 0
\]

\hypertarget{result-2}{%
\subsubsection{13.4 Result}\label{result-2}}

The total number of internal free parameters is:

\[
\dim(\text{theory}) = 0
\]

\hypertarget{interpretation-4}{%
\subsubsection{13.5 Interpretation}\label{interpretation-4}}

All primary structures arise from:

\begin{itemize}
\tightlist
\item
  discrete topology
\item
  symmetry constraints
\item
  locality and minimality
\end{itemize}

No continuous parameter is introduced prior to phenomenological
matching.

Thus:

\begin{quote}
The \(\ell^1\) causal lattice defines a \textbf{zero-parameter
structural framework} at the level of its internal construction.
\end{quote}

\hypertarget{relation-to-physical-constants}{%
\subsubsection{13.6 Relation to Physical
Constants}\label{relation-to-physical-constants}}

Observed physical constants enter in two ways:

\hypertarget{structurally-determined}{%
\paragraph{(1) Structurally determined}\label{structurally-determined}}

Examples:

\begin{itemize}
\tightlist
\item
  transport amplitude (\(3/\pi\))
\item
  gauge decomposition
\item
  generation count
\end{itemize}

These arise from zero-parameter construction.

\hypertarget{phenomenologically-anchored}{%
\paragraph{(2) Phenomenologically
anchored}\label{phenomenologically-anchored}}

Examples:

\begin{itemize}
\tightlist
\item
  cosmological constant (\(\Lambda\))
\item
  fermion mass spectrum
\end{itemize}

These correspond to:

\begin{quote}
quantities not yet derived from the discrete structure
\end{quote}

and are explicitly identified as open problems.

\hypertarget{consequence}{%
\subsubsection{13.7 Consequence}\label{consequence}}

The framework implies:

\begin{quote}
any successful geometric or symmetry extension must preserve zero
internal degrees of freedom
\end{quote}

or identify precisely where additional topological structure enters.

\hypertarget{falsifiability}{%
\subsubsection{13.8 Falsifiability}\label{falsifiability}}

The zero-parameter nature introduces a strong constraint:

\begin{quote}
any alternative construction satisfying the same structural axioms must
reproduce the same outputs.
\end{quote}

Thus, the framework is falsified if:

\begin{itemize}
\tightlist
\item
  an admissible operator exists with amplitude \(\neq 3/\pi\), or
\item
  an admissible symmetry structure differs from
  \(SU(3) \times SU(2) \times U(1)\), or
\item
  a consistent \(\ell^1\) construction produces additional free
  parameters
\end{itemize}

\hypertarget{summary}{%
\subsubsection{13.9 Summary}\label{summary}}

The \(\ell^1\) causal lattice is not:

\begin{itemize}
\tightlist
\item
  a parameterized model,
\item
  nor a family of theories
\end{itemize}

but:

\begin{quote}
a single constrained construction with no internal degrees of freedom.
The only remaining freedom lies in the existence of the initial
inconsistency itself, which is not a parameter of the theory but the
condition under which any theory becomes possible.
\end{quote}

\hypertarget{on-the-scope-of-uniqueness-claims}{%
\subsubsection{13.10 On the Scope of Uniqueness
Claims}\label{on-the-scope-of-uniqueness-claims}}

The uniqueness results established in Sections 11--13 are conditional on
the specified constraint set:

\begin{itemize}
\tightlist
\item
  locality (single-step transport)
\item
  linearity
\item
  symmetry (shift invariance, parity)
\item
  minimal topology (2-simplex)
\end{itemize}

A potential objection is that alternative admissible constructions may
exist outside this constraint set.

This objection is valid in principle. However:

\begin{quote}
any such construction must explicitly identify which constraint is
relaxed and demonstrate that the resulting structure preserves the
intended physical invariants.
\end{quote}

In particular, alternatives must account for:

\begin{itemize}
\tightlist
\item
  stability under composition
\item
  bounded transport
\item
  compatibility with discrete topology
\end{itemize}

Absent such a construction, the classification remains complete within
the stated domain.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{open-problems}{%
\subsection{14. Open Problems}\label{open-problems}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Derive the cosmological constant quantitatively.} The
  subdivision overshoot mechanism predicts \(\Lambda > 0\) and explains
  its smallness qualitatively. A quantitative prediction requires a
  theory of the average overshoot rate.
\item
  \textbf{Derive the dark matter mass spectrum.} The non-resonant defect
  mechanism predicts massive, invisible, gravitating, stable particles.
  The mass distribution requires understanding which non-resonant
  configurations are dynamically accessible.
\item
  \textbf{Formalize the coarse-graining steps.} The transitions
  \(\ell^1 \to \ell^2\) (Parseval) and \(\ell^2 \to \ell^\infty\) (law
  of large numbers) are identified but not rigorously proven to be the
  unique coarse-graining maps preserving the relevant physical
  structure.
\item
  \textbf{Derive fermion masses and mixing.} The three generations are
  forced (Paper 009), but the mass hierarchy within and between
  generations remains underived.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{conclusion}{%
\subsection{15. Conclusion}\label{conclusion}}

This work proposes that a substantial portion of known physical
structure can be derived from a single organizing principle: monotone
correction of local inconsistency under the \(\ell^1\) coboundary norm.

Starting from this principle and a minimal discrete seed, the framework
constructs a sequence of constrained structures, including:

\begin{itemize}
\tightlist
\item
  a discrete \(N = 3\) lattice and associated Fourier structure,
\item
  Hilbert space behavior via \(\ell^1 \to \ell^2\) correspondence,
\item
  a constrained transport amplitude and associated unitarity deviation,
\item
  the Standard Model gauge group \(SU(3) \times SU(2) \times U(1)\),
\item
  three fermion generations and electroweak structure,
\item
  discrete geometric tiling and curvature via Regge defects,
\item
  the Planck scale as a defect threshold,
\item
  and a hierarchy of effective descriptions arising from norm
  transitions.
\end{itemize}

Within this framework, several longstanding problems are reinterpreted:

\begin{itemize}
\tightlist
\item
  The hierarchy between gravitational and electroweak scales appears as
  a statement about lattice extent rather than parameter fine-tuning.
\item
  The cosmological constant reflects residual imbalance in a locally
  driven subdivision process rather than vacuum energy divergence.
\item
  Dark matter corresponds to defect configurations that do not couple to
  the Standard Model gauge structure.
\end{itemize}

The framework does not yet provide quantitative derivations of the
cosmological constant, a global dark matter mass spectrum, or fermion
mass hierarchies. These remain open problems.

The central claim is therefore not that all physical quantities have
been derived, but that:

\begin{quote}
a large fraction of known structure can be organized as the consequence
of a single constrained discrete mechanism, with remaining parameters
isolated and clearly identified.
\end{quote}

If correct, this shifts the role of unification:

\begin{quote}
not as the merger of independent theories, but as the identification of
a common discrete substrate from which multiple effective descriptions
emerge.
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{references}{%
\subsection{References}\label{references}}

{[}1{]} T. Regge, ``General relativity without coordinates,'' \emph{Il
Nuovo Cimento} \textbf{19}, 558 (1961).

{[}2{]} J. Maldacena, ``The large-\(N\) limit of superconformal field
theories and supergravity,'' \emph{Int. J. Theor. Phys.} \textbf{38},
1113 (1999).

{[}3{]} C. Rovelli, \emph{Quantum Gravity} (Cambridge University Press,
2004).

{[}4{]} J. Polchinski, \emph{String Theory}, Vols. 1--2 (Cambridge
University Press, 1998).

{[}5{]} R. D. Sorkin, ``Causal sets: Discrete gravity,'' in
\emph{Lectures on Quantum Gravity}, eds.~A. Gomberoff and D. Marolf
(Springer, 2005).

{[}6{]} Particle Data Group (R. L. Workman et al.), ``Review of Particle
Physics,'' \emph{PTEP} \textbf{2022}, 083C01 (2022). Updated 2024.

{[}7{]} Planck Collaboration (N. Aghanim et al.), ``Planck 2018 results.
VI. Cosmological parameters,'' \emph{Astron. Astrophys.} \textbf{641},
A6 (2020).

{[}8{]} J. H. Carroll, ``Paper 000: Projection Obstruction Theory:
Retraction Nonlinearity, l1 Rigidity, and Density Scaling'' (2026).

{[}9{]} J. H. Carroll, ``Paper 001: The Free \(\ell^1\) Seminorm on
Banach Presheaf Coboundaries,'' (2026).

{[}10{]} J. H. Carroll, ``Paper 002: Coordinate-Wise Additivity and the
\(\ell^1\) Norm on Finite Graph Cochains'' (2026).

{[}10{]} J. H. Carroll, ``Paper 003: Hodge Structure and Gauge
Equivalence of \(\ell^1\) Defect Fields'' (2026).

{[}11{]} J. H. Carroll, ``Paper 004: Universal Obstruction Theory - The
\(\ell^1\) Topological Framework'' (2026).

{[}12{]} J. H. Carroll, ``Paper 005: Autopoietic Cohomology: Iterative
Obstruction Repair on Causal Posets,'' (2026).

{[}13{]} J. H. Carroll, ``Paper 006: The Unification Framework:
\(\hbar\), \(c\), and \(G\) as Structural Scaling Parameters for
\(\ell^1\) Defect Systems'' (2026).

{[}14{]} J. H. Carroll, ``Paper 007: Topological Constraint on the
Electromagnetic Coupling Scale from a Discrete \(\ell^1\) Obstruction''
(2026).

{[}15{]} J. H. Carroll, ``Paper 008: Constraint-Based Selection of Gauge
Structure from \(\ell^1\) Topology'' (2026).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\end{document}
