% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{amsart}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Global \textbackslash ell\^{}1 Rigidity on Finite Graph Cochains},
  pdfauthor={Jeremy H. Carroll},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\title{Global \(\ell^1\) Rigidity on Finite Graph Cochains}
\author{Jeremy H. Carroll}
\date{}

\begin{document}
\begin{abstract}
This paper proves a classification theorem for seminorms on graph
cochains, showing that structural independence and symmetry constraints
force \(\ell^1\) geometry. Any seminorm on \(C^1(G,\mathbb{R}^k)\) that connects
independent measurements---acting locally across edges and additively
across disjoint coordinates---under permutation symmetry must be a
scalar multiple of the \(\ell^1\) norm. This establishes that the defect
field is isometrically equivalent to \(L^1(E \times \{1,\dots,k\})\)
under the counting measure. The resulting \(\ell^1 \otimes \ell^1\)
metric demonstrates why total variation, graph cuts, and projection
obstruction correlation measures inevitably restrict to sparse
\(\ell^1\) aggregation.
\end{abstract}
\maketitle

\hypertarget{introduction}{%
\subsection{1. Introduction}\label{introduction}}

This paper completes a trilogy studying projection obstructions on
presheaf coboundaries:

\begin{itemize}
\tightlist
\item
  \textbf{Paper 000} {[}6{]} establishes that projecting dynamics onto a
  product-state manifold creates a geometric curvature obstruction over
  correlation defects.
\item
  \textbf{Paper 001} {[}7{]} demonstrates that aggregation across edges
  must follow an \(\ell^1\) rule.
\end{itemize}

A central mathematical vulnerability remains: why should the local
diagnostic norm itself, operating on the \(k\)-dimensional measurement
channels (the stalks), assume an \(\ell^1\) geometry?

Rather than engineering axioms to manufacture an \(\ell^1\) result, we
derive the full diagnostic form strictly from fundamental structural
physics. When forced to respect both axes of separability---edges and
coordinates---the diagnosis becomes globally rigid.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{ell1-characterization-on-graph-cochains}{%
\subsection{\texorpdfstring{2. \(\ell^1\) Characterization on Graph
Cochains}{2. \textbackslash ell\^{}1 Characterization on Graph Cochains}}\label{ell1-characterization-on-graph-cochains}}

\textbf{Theorem 2.1.} Let \(G=(V,E)\) be a finite directed graph and
consider the cochain space
\[ C^1(G,\mathbb{R}^k)=\bigoplus_{e\in E}\mathbb{R}^k. \]

Let \(\nu:C^1(G,\mathbb{R}^k)\to\mathbb{R}_{\ge0}\) be a seminorm
satisfying:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Edge locality} \[ \nu(\delta)=\sum_{e\in E}\nu_e(\delta_e) \]
\item
  \textbf{Coordinate symmetry}
  \[ \nu_e(x_{\sigma(1)},\dots,x_{\sigma(k)}) = \nu_e(x_1,\dots,x_k) \]
  for every permutation \(\sigma\).
\item
  \textbf{Coordinate additivity} If \(x,y\in\mathbb{R}^k\) have disjoint
  coordinate support
  (\(\mathrm{supp}(x) \cap \mathrm{supp}(y) = \varnothing\) where
  \(\mathrm{supp}(x) = \{i : x_i \neq 0\}\)), then
  \[ \nu_e(x+y)=\nu_e(x)+\nu_e(y). \]
\item
  \textbf{Graph invariance} For any graph isomorphism
  \(\sigma:G\to G'\), \[ \nu_{G'}(\sigma_*\delta)=\nu_G(\delta). \]
\end{enumerate}

Then there exists \(\alpha > 0\) such that
\[ \nu(\delta)=\alpha\sum_{e\in E}\|\delta_e\|_1. \]

\textbf{Proof.}

By edge locality it suffices to classify the seminorm \(\nu_e\) on
\(\mathbb{R}^k\).

Let \(x=(x_1,\dots,x_k)\). Decompose \[ x=\sum_{i=1}^k x_i \hat{e}_i. \]

The vectors \(x_i \hat{e}_i\) have disjoint coordinate support, so
coordinate additivity gives
\[ \nu_e(x)=\sum_{i=1}^k\nu_e(x_i \hat{e}_i). \]

Since \(\nu_e\) is a seminorm, it is absolutely homogeneous, providing
\[ \nu_e(x_i \hat{e}_i)=|x_i|\nu_e(\hat{e}_i). \]

By coordinate symmetry, \[ \nu_e(\hat{e}_i)=w_e \] for all \(i\).

Hence \[ \nu_e(x)=w_e\sum_{i=1}^k |x_i|=w_e\|x\|_1. \]

Thus across the entire graph,
\[ \nu(\delta)=\sum_{e\in E} w_e\|\delta_e\|_1. \]

Finally, graph invariance implies all intrinsic weights coincide:
\[ w_e=\alpha. \]

Therefore \[ \nu(\delta)=\alpha\sum_{e\in E}\|\delta_e\|_1. \]
\(\square\)

\hypertarget{corollary-2.2-discrete-l1-structure-of-the-defect-space}{%
\subsection{\texorpdfstring{Corollary 2.2 (Discrete \(L^1\) Structure of
the Defect
Space)}{Corollary 2.2 (Discrete L\^{}1 Structure of the Defect Space)}}\label{corollary-2.2-discrete-l1-structure-of-the-defect-space}}

Under the hypotheses of Theorem 2.1, the cochain space equipped with the
diagnostic seminorm satisfies

\[
(C^1(G,\mathbb{R}^k),\nu)
\cong
L^1(E\times\{1,\dots,k\})
\]

with respect to the counting measure.

\textbf{Proof.}

By Theorem 2.1,

\[
\nu(\delta)
=
\alpha\sum_{e\in E}\|\delta_e\|_1.
\]

Writing \(\delta_{e,i}\) for the \(i\)-th coordinate of \(\delta_e\), we
obtain

\[
\nu(\delta)
=
\alpha\sum_{e\in E}\sum_{i=1}^k |\delta_{e,i}|.
\]

Define

\[
f:E\times\{1,\dots,k\}\to\mathbb{R}
\]

by

\[
f(e,i)=\delta_{e,i}.
\]

Then

\[
\nu(\delta)
=
\alpha\|f\|_{L^1}.
\]

Thus the diagnostic space is isometrically equivalent to\\
\(L^1(E\times\{1,\dots,k\})\) under the counting measure.\\
\(\square\)

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{structural-consequences}{%
\subsection{3. Structural Consequences}\label{structural-consequences}}

\hypertarget{coarea-duality-k-1}{%
\subsubsection{\texorpdfstring{3.1 Coarea Duality
(\(k = 1\))}{3.1 Coarea Duality (k = 1)}}\label{coarea-duality-k-1}}

For scalar signals, the classified norm has a distinguished topological
property.

\textbf{Fact 3.1 (Discrete Coarea Formula} {[}2, 3{]}\textbf{).} For
\(v: V \to \mathbb{R}\):
\[ \text{TV}_G(v) = \int_{-\infty}^{\infty} |\partial V_t| \, dt \]
where \(V_t = \{m : v_m > t\}\) and \(|\partial V_t|\) counts edges
crossing the boundary.

When boundary conditions force a binary partition
(\(v_a = 1, v_b = 0\)), \(\ell^1\) minimization recovers the minimum
\(s\)-\(t\) cut via max-flow/min-cut {[}5{]}.

\textbf{Remark 3.2.} For \(k > 1\), the total variation
\(\sum_e \|v_n - v_m\|_1\) decomposes (by Theorem 2.1) as a sum of \(k\)
independent scalar total variations, each individually satisfying the
coarea formula.

\hypertarget{the-geometry-of-the-defect-field}{%
\subsubsection{3.2 The Geometry of the Defect
Field}\label{the-geometry-of-the-defect-field}}

The explicit inclusion of \textbf{Coordinate Additivity} formally
captures the physical independence of disparate measurement channels.
Permutation symmetry and seminorm subadditivity alone permit
pathological metrics like the \(\ell^\infty\) maximum norm or the
Euclidean \(\ell^2\) norm. The strict structural requirement that
physically isolated defect channels---the orthogonal coordinates---must
accumulate their structural signatures linearly thereby forcing the
\(\ell^1\) stalk geometry.

Thus, the unique seminorm compatible with the stated structural
principles of locality, measurement independence, and symmetry is the
\(\ell^1\) metric: \(\ell^1 \otimes \ell^1\).

Conceptually, our theorem establishes that the defect space restricts
to: \[ C^1(G,\mathbb{R}^k) \cong L^1(E \times \{1,\dots,k\}) \] whenever
the diagnostic respects independent components. This serves as a finite
combinatorial analogue of Kakutani's (1941) concrete characterization of
abstract \(L\)-spaces {[}1{]}. Additivity across independent structural
components forces the \(\ell^1\) geometry over the counting measure.
Defect fields thus behave strictly as discrete 1-forms whose absolute
total variation tracks graph cuts and sparse obstructions {[}3{]}.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{acknowledgments}{%
\subsection{Acknowledgments}\label{acknowledgments}}

No external funding. No conflicts of interest.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{references}{%
\subsection{References}\label{references}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Kakutani, S. (1941). \emph{Concrete representation of abstract
  \((L)\)-spaces and the mean ergodic theorem}. Annals of Mathematics,
  42(2), 523-537.
\item
  Fleming, W. H., \& Rishel, R. (1960). \emph{An integral formula for
  total gradient variation}. Archiv der Mathematik, 11(1), 218-222.
\item
  Chambolle, A., et al.~(2010). \emph{An introduction to total variation
  for image analysis.} In: Theoretical Foundations and Numerical Methods
  for Sparse Recovery. De Gruyter.
\item
  Rockafellar, R. T. (1970). \emph{Convex Analysis.} Princeton
  University Press.
\item
  Ford, L. R., \& Fulkerson, D. R. (1956). \emph{Maximal flow through a
  network.} Canadian Journal of Mathematics, 8, 399--404.
\item
  Carroll, J. H. (2025). \emph{Single Obstruction Theory: Projection
  Nonlinearity, \(\ell^1\) Rigidity, and Density Scaling.} Zenodo.
  https://doi.org/10.5281/zenodo.18896777
\item
  Carroll, J. H. (2026). \emph{The Free \(\ell^1\) Seminorm on Banach
  Presheaf Coboundaries.} Zenodo.
  https://doi.org/10.5281/zenodo.18897081
\end{enumerate}

\end{document}
